### 2021-12-02 cn/publications.tex

% AGU style: https://publications.agu.org/agu-grammar-and-style-guide/
\newcommand{\Revision}{\textit{正在审稿}}
\newcommand{\CS}{*} % corresponding author
\newcommand{\CF}{\textsuperscript{\#}} % co-first author

\section*{已发表论文}
\CS 通讯作者，\CF 共同一作

\begin{etaremune}
\item Yao, J., \Tian, Sun, L., \& Wen, L. (2021).
	Comment on ``Origin of temporal changes of inner-core seismic waves'' by Yang and Song (2020).
	\textit{Earth and Planetary Science Letters}, \textit{553}, 116640.
	\DOI{10.1016/j.epsl.2020.116640}
\item Wei, S. S., Shearer, P. M., Lithgow-Bertelloni, C., Stixrude, L., \& \Tian\ (2020).
	Oceanic plateau of the Hawaiian mantle plume head subducted to the uppermost lower mantle.
	\textit{Science}, \textit{370}, 983--987.
	\DOI{10.1126/science.abd0312}
\item \Tian\CS, Lv, M., Wei, S. S., Dorfman, S. M. \& Shearer, P. M. (2020).
	Global variations of Earth's 520- and 560-km discontinuities.
	\textit{Earth and Planetary Science Letters}, \textit{552}, 116600. \\
	\DOI{10.1016/j.epsl.2020.116600}
\item
    Wessel, P., Luis, J., Uieda, L., Scharroo, R., Wobbe, F., Smith, W. H. F., \& \Tian\ (2019).
    The Generic Mapping Tools Version 6.
    \textit{Geochemistry, Geophysics, Geosystems}, \textit{20}(11), 5556--5564.
    \DOI{10.1029/2019GC008515}
\item
    Yao, J., \Tian, Sun, L., \& Wen, L. (2019).
    Temporal change of seismic Earth's inner core phases: inner core differential rotation or temporal change of inner core surface?.
    \textit{Journal of Geophysical Research: Solid Earth}, \textit{124}(7), 6720--6736.
    \DOI{10.1029/2019JB017532}
\item
    Fan, W., Wei, S.S.,\Tian, McGuire J.J., \& Wiens D.A. (2019).
    Complex and diverse rupture processes of the 2018 Mw 8.2 and Mw 7.9 Tonga-Fiji deep earthquakes.
    \textit{Geophysical Research Letters}, \textit{46}(5), 2434--2448.
    \DOI{10.1029/2018GL080997}
\item
    Yao, J., \Tian\CF, Lu, Z., Sun, L., \& Wen, L. (2018).
    Triggered seismicity after North Korea's 3 September 2017 nuclear test.
    \textit{Seismological Research Letters}, \textit{89}(6), 2085--2093.
    \DOI{10.1785/0220180135}
\item
    Yao, J., \Tian\CF, Sun, L., \& Wen, L. (2018).
	Source characteristics of North Korea's 3 September 2017 nuclear test.
    \textit{Seismological Research Letters}, \textit{89}(6), 2078--2084.
    \DOI{10.1785/0220180134}
\item
    \Tian\CS, Yao, J., \& Wen, L. (2018).
    Collapse and earthquake swarm after North Korea's 3 September 2017 nuclear test.
    \textit{Geophysical Research Letters}, \textit{45}(9), 3976--3983.
    \DOI{10.1029/2018GL077649}
\item
    Wen, L., \Tian, \& Yao, J. (2018).
    Seismic structure and dynamic process of the Earth's inner core and its boundary.
    \textit{Chinese Journal of Geophysics}, \textit{61}(3), 803--818.
    \DOI{10.6038/cjg2018L0500} [in Chinese]
\item
    \Tian, \& Wen, L. (2017).
    Seismological evidence for a localized mushy zone at the Earth's inner core boundary.
    \textit{Nature communications}, 8, 165.
    \DOI{10.1038/s41467-017-00229-9}
\item
    Chen, X., \Tian, \& Wen, L. (2015).
    Microseismic sources during hurricane sandy.
    \textit{Journal of Geophysical Research: Solid Earth}, \textit{120}(9), 6386--6403.
    \DOI{10.1002/2015JB012282}
\item Zhang, M., \Tian, \& Wen, L. (2014).
    A new method for earthquake depth determination: stacking multiple-station autocorrelograms.
    \textit{Geophysical Journal International}, \textit{197}(2), 1107--1116.
    \DOI{10.1093/gji/ggu044}
\end{etaremune}

%\subsection*{尚未发表论文（审稿中/修改中）}
%\begin{etaremune}
%\end{etaremune}