### 2021-12-02 en/activities.tex

\section*{Professional Societies \& Activities}

\begin{itemize}
\item Member of the \href{https://sites.agu.org/}{American Geophysical Union (AGU)} (2012 -- present)
\item Peer-reviewer of scientific journals:
      \textit{Geophysical Research Letters},
      \textit{Seismological Research Letters},
      \textit{Review of Scientific Instruments},
      \textit{Journal of Open Source Software},
      \textit{Results in Geophysical Sciences}
\item Founder of the \href{https://blog.seisman.info}{SeisMan blog} (since 2013),
      \href{http://gmt-china.org/}{GMT China Community} (since 2016)
      and \href{https://seismo-learn.org/}{seismo-learn} (since 2020)
\item Core developer of the \href{https://github.com/GenericMappingTools/gmt}{Generic Mapping Tools (GMT)} and \href{https://github.com/GenericMappingTools/pygmt}{PyGMT} (2018 -- present)
\item Research assistant and database manager for \href{http://chinageorefmodel.org/}{China Seismological Reference Model} (2016 -- 2018)
\item Judge for the Outstanding Student Paper Award, AGU Fall Meeting (2018 -- 2020)
\item Instructor, the UNAVCO Short Course ``The Generic Mapping Tools for Geodesy'' (2019 -- 2021)
\item Instructor, Workshop SCIWS4: \href{https://www.agu.org/Events/SCIWS4-Generic-Mapping-Tools}{``Become a Generic Mapping Tools Contributor Even If You Can't Code''}, 2019 AGU Fall Meeting (2019)
\end{itemize}