### 2021-12-02 en/talks.tex

\section*{Talks}

\begin{etaremune}
\item
	Tectonic-Geophysics Young Scholars Research Symposium.
	Nanjing University.
	Jan. 7, 2021.
\item
	Department of Earth and Space Sciences, Southern University of Science and Techbology.
	Nov. 27, 2020.
	\invited.
\item
    2nd Annual Earth and Environmental Sciences Student Research Symposium.
    Department of Earth and Environmental Sciences, Michigan State University.
    Feb. 23, 2019.
\item
    Institute of Geology and Geophysics, Chinese Academy of Sciences.
    Jun. 15, 2018.
    \invited
\item
    Institute of Earthquake Forcasting, China Earthquake Administration.
    Jun. 14, 2018.
\item
    2017 Annual Meeting of Chinese Geoscience Union (CGU).
    Oct. 17, 2017.
    \invited
\item
    Workshop on Analysis and Applications of Crustal Deformation Data.
    Hubei Earthquake Administration.
    Sep. 21, 2016.
    \invited
\item
    China Earthquake Networks Center.
    Jun. 30, 2016.
    \invited
\end{etaremune}